The Impact of Accretion Disk Winds on the Optical Spectra of Cataclysmic Variables

### 2015-03-24 old.tex
starting to address referee comments

\title{The Impact of Accretion Disk Winds on the Optical Spectra of Cataclysmic Variables}
\author{J. H. Matthews, C. Knigge, K. S. Long, S. A. Sim \& N. Higginbottom}

@@

\begin{abstract}
Many high-state non-magnetic cataclysmic variables (CVs) exhibit
blue-shifted absorption or P-Cygni profiles associated with ultraviolet
(UV) resonance lines, such as C~{\sc iv}~$\lambda1550$. These features
imply the existence of powerful accretion disk winds in
CVs. Here, we use an improved version of our Monte Carlo ionization
and radiative transfer code to investigate whether disk wind models that
produce realistic UV line profiles are also likely to generate
observationally significant recombination line and continuum emission
in the {\em optical} waveband. 
We also test whether outflows may be responsible for the single-peaked 
emission line profiles often seen
in high-state CVs and for the weakness of the Balmer absorption edge
(relative to simple models of optically thick accretion disks).
We find that a standard disk wind model that is successful in
reproducing the UV spectra of CVs also leaves a noticeable imprint on
the optical spectrum, particularly for systems viewed at high
inclination. The strongest optical wind-formed recombination lines are \ha\ and
\heiiopt, although the outflow also produces emission in the Ly$\alpha$
and \heiiuv\ UV lines. However, all of the optical lines produced by
this default wind model are double-peaked, and the effect of the wind
on the Balmer jump is modest. We demonstrate that a higher-density outflow model 
displays reasonable verisimilitude with the optical spectrum of RW Trianguli,
producing all the observed lines. This model
produces a recombination continuum that is strong enough to fill in the Balmer jump,
but it also predicts stronger-than-observed collisionally excited UV emission lines. 
The wind in our revised model also alters the spectral shape of the continuum, 
and displays a strong inclination dependence in its emission properties. 
No single-peaked emission is seen, although we observe a 
narrowing of the double-peaked emission lines from the base of the wind.
We show that even denser models can produce a single-peaked \ha\ line.
In this model, the single-peaked line profile is due to the line emission originating from 
$\sim10^{11}$~cm above the disk plane where rotational velocities are small, 
and is not due to the radial velocity shear mechanism that
has previously been proposed. On the basis of our results, we suggest that winds can 
modify, and perhaps even dominate, the line and continuum emission
from CVs.
\end{abstract}

>>>>

\title{The Impact of Accretion Disk Winds on the Optical Spectra of Cataclysmic Variables}

\author[Matthews et al.]{J.~H.~Matthews,$^1$ C.~Knigge,$^1$ K.~S.~Long,$^2$ S.~A.~Sim,$^3$ and N.~Higginbottom$^4$
\medskip  
\\$^1$School of Physics and Astronomy, University of Southampton, Highfield, Southampton, SO17 1BJ, UK
\\$^2$Space Telescope Science Institute, 3700 San Martin Drive, Baltimore, MD 21218, USA
\\$^3$School of Mathematics and Physics, Queens University Belfast, University Road, Belfast, BT7 1NN, Northern Ireland, UK
\\$^4$Department of Astronomy, University of Las Vegas, Las Vegas, NV 89119, USA}

%%\author{J. H. Matthews, C. Knigge, K. S. Long, S. A. Sim \& N. Higginbottom}

@@

\begin{abstract}
% Many high-state non-magnetic cataclysmic variables (CVs) exhibit
% blue-shifted absorption or P-Cygni profiles associated with ultraviolet
% (UV) resonance lines, such as C~{\sc iv}~$\lambda1550$. These features
% imply the existence of powerful accretion disk winds in
% CVs. Here, we use an improved version of our Monte Carlo ionization
% and radiative transfer code to investigate whether disk wind models that
% produce realistic UV line profiles are also likely to generate
% observationally significant recombination line and continuum emission
% in the {\em optical} waveband. 
% We also test whether outflows may be responsible for the single-peaked 
% emission line profiles often seen
% in high-state CVs and for the weakness of the Balmer absorption edge
% (relative to simple models of optically thick accretion disks).
% We find that a standard disk wind model that is successful in
% reproducing the UV spectra of CVs also leaves a noticeable imprint on
% the optical spectrum, particularly for systems viewed at high
% inclination. The strongest optical wind-formed recombination lines are \ha\ and
% \heiiopt, although the outflow also produces emission in the Ly$\alpha$
% and \heiiuv\ UV lines. However, all of the optical lines produced by
% this default wind model are double-peaked, and the effect of the wind
% on the Balmer jump is modest. We demonstrate that a higher-density outflow model 
% displays reasonable verisimilitude with the optical spectrum of RW Trianguli,
% producing all the observed lines. This model
% produces a recombination continuum that is strong enough to fill in the Balmer jump,
% but it also predicts stronger-than-observed collisionally excited UV emission lines. 
% The wind in our revised model also alters the spectral shape of the continuum, 
% and displays a strong inclination dependence in its emission properties. 
% No single-peaked emission is seen, although we observe a 
% narrowing of the double-peaked emission lines from the base of the wind.
% We show that even denser models can produce a single-peaked \ha\ line.
% In this model, the single-peaked line profile is due to the line emission originating from 
% $\sim10^{11}$~cm above the disk plane where rotational velocities are small, 
% and is not due to the radial velocity shear mechanism that
% has previously been proposed. On the basis of our results, we suggest that winds can 
% modify, and perhaps even dominate, the line and continuum emission
% from CVs.
Many high-state non-magnetic cataclysmic variables (CVs) exhibit
blue-shifted absorption or P-Cygni profiles associated with ultraviolet
(UV) resonance lines. These features imply the existence of 
powerful accretion disk winds in CVs. Here, we use our Monte Carlo ionization
and radiative transfer code to investigate whether disk wind models that
produce realistic UV line profiles are also likely to generate
observationally significant recombination line and continuum emission
in the {\em optical} waveband. We also test whether outflows may be responsible for 
the single-peaked emission line profiles often seen in high-state CVs and for the weakness
of the Balmer absorption edge (relative to simple models of optically thick accretion disks).
We find that a standard disk wind model that is successful in
reproducing the UV spectra of CVs also leaves a noticeable imprint on
the optical spectrum, particularly for systems viewed at high
inclination. The strongest optical wind-formed recombination lines are \ha\ and
\heiiopt. We demonstrate that a higher-density outflow model produces 
all the expected H and He lines and produces a recombination continuum that can 
fill in the Balmer jump at high inclinations.
This model displays reasonable verisimilitude with the optical spectrum of RW Trianguli. No single-peaked emission is seen, although we observe a 
narrowing of the double-peaked emission lines from the base of the wind.
Finally, we show that even denser models can produce a single-peaked \ha\ line.
On the basis of our results, we suggest that winds can 
modify, and perhaps even dominate, the line and continuum emission
from CVs.
\end{abstract}

\begin{keywords}
{\sl (stars:)} novae, cataclysmic variables - accretion, accretion discs - stars: winds, outflows - line: profiles - radiative transfer - methods: numerical 
\end{keywords}
